---
title: "Potential lunar-modulated rhythm of magnetic sensing in the Oyster toadfish"
output: html_document
date: "2024-01-25"
---

################################################################################
Load libraries
```{r}
library(RColorBrewer)
library(circular)
library(ggplot2)
library(ggpubr)
```


################################################################################
Color
```{r}
# Hexadecimal color specification 
brewer.pal(n = 8, name = "BuGn")
# "#F7FCFD" "#E5F5F9" "#CCECE6" "#99D8C9" "#66C2A4" "#41AE76" "#238B45" "#005824"
display.brewer.pal(7,"BrBG")
```



################################################################################
Lunar Day ~ Den Choice Plotting and Statistical Analysis
Total n= 98
Control n= 51
Magnet n= 47

```{r}
control_lunar_data <- read.csv("2024_01_25 Control Lunar Day Data.csv", sep = ",", header = TRUE)
control_lunar_data
control_choice <- lunar_data$Control_Lunar_Day_Degree
control_choice_circ <- circular(control_choice, units = "degrees", template = "geographics", modulo = "2pi") 
mean.circular(control_choice_circ)
#354.5702
rayleigh.test(control_choice_circ)
#Test Statistic:  0.1681  
#P-value:  0.2366 
```


```{r}
plot.circular(control_choice_circ, col= "#D8B365", stack = TRUE, bins = 360, sep = .07, shrink = 1.5)
```


```{r}
magnet_lunar_data <- read.csv("2024_01_25 Magnet Lunar Day Data.csv", sep = ",", header = TRUE)
magnet_lunar_data
magnet_choice <- magnet_lunar_data$Magnet_Lunar_Day_Degree
magnet_choice_circ <- circular(magnet_choice, units = "degrees", template = "geographics", modulo = "2pi") 
mean.circular(magnet_choice_circ)
#182.0084
rayleigh.test(magnet_choice_circ)
#Test Statistic:  0.2152  
#P-value:  0.1133
```

```{r}
plot.circular(magnet_choice_circ, col= "#5AB4AC", stack = TRUE, bins = 360, sep = .07, shrink = 1.5)
```

Statistical testing
```{r}
watson.two.test(control_choice_circ, magnet_choice_circ)
#Test Statistic: 0.1912
#0.01 < P-value < 0.05  

watson.wheeler.test(list(control_choice_circ, magnet_choice_circ))
#W = 6.3286, df = 2, p-value = 0.04224
```


################################################################################
Moon Illumination and Fish Length ~ Den Choice Logistic Regression Model

```{r}
regression_data <- read.csv("2024_01_25 Regression Data.csv", sep = ",", header = TRUE)
moon_model <- glm(regression_data$Morning_DV ~ regression_data$Moon_Illumination, family = binomial)
summary(moon_model)
#                               Estimate Std. Error z value Pr(>|z|)  
#(Intercept)                        -0.770054   0.356953  -2.157   0.0310 *
#regression_data$Moon_Illumination  0.014111   0.005903   2.390   0.0168 *

#calculating CI
confint(moon_model)
#                                     2.5 %      97.5 %
#(Intercept)                        -1.495487620 -0.08627778
#regression_data$Moon_Illumination  0.002767395  0.02602998

#calculating odds ratio
moon_model$coefficients
exp(moon_model$coefficients)
#(Intercept) regression_data$Moon_Illumination 
#0.462988                          1.014211

```


```{r}
regression_plot <- ggplot(regression_data, aes(x=Moon_Illumination, y = Morning_DV)) + geom_point(alpha=.4, size=1.8, position=position_jitter(width=.04,height=.04)) + stat_smooth(method = "glm", color = "black", se=TRUE, method.args = list(family="binomial")) + theme_classic() + labs (x="Moon Illumination (%)", y="Probability of Choosing Magnet Den") + theme(axis.text = element_text(size=12), axis.title = element_text(size=14))
regression_plot
ggsave("lunarregression.pdf", width=6, height=4, dpi=300)
```

```{r}
length_model <- glm(regression_data$Morning_DV ~ regression_data$Length, family = binomial)
summary(length_model)
#                       Estimate Std. Error z value Pr(>|z|)
#(Intercept)             1.31324    1.47604   0.890    0.374
#regression_data$Length -0.05198    0.05449  -0.954    0.340
```


################################################################################
Den Choice in the Morning Plotting and Statistical Analysis

```{r}
choice_data <- read.csv("2024_01_25 Den Choice.csv", sep = ",", header = TRUE)
morning_choice_barplot <- ggplot(choice_data, aes(x= Lunar_Phase, y=Lunar_Phase_Morning_Dec, fill=Lunar_Phase_Den)) + geom_col(position = 'dodge') + theme(axis.text = element_text(size=12), axis.title = element_text(size=14)) + theme_classic() +
  labs (x="Lunar Phase", y="Number of Fish")  + scale_x_discrete(limits=c("New Moon","Full Moon")) + ylim(0,20) + scale_fill_manual(values=c("#D8B365", "#5AB4AC", "#5AB4AC"))
morning_choice_barplot
ggsave("morning_choice.pdf", width=6, height=4, dpi=300)
```
Full Moon (n=25): control n=8, magnet n=17
```{r}
binom.test(8, 25, p = 0.5)
#number of successes = 8, number of trials = 25, p-value = 0.1078
```

New Moon (n=27): control n=20, magnet n=7
```{r}
binom.test(20, 27, p = 0.5)
#number of successes = 20, number of trials = 27, p-value = 0.01916
```

Control Choices (N=28): new moon (n=20) and full moon (n=8)
```{r}
binom.test(20, 28, p = 0.5)
#number of successes = 20, number of trials = 28, p-value = 0.0357
```

Magnet Choices (N=24): new moon (n=7) and full moon (n=17)
```{r}
binom.test(17, 24, p = 0.5)
#number of successes = 17, number of trials = 24, p-value = 0.06391
```

################################################################################
Den Preference Plotting and Statistical Analysis

```{r}
nm_pref_data <- read.csv("2024_01_25 New Moon Den Preference.csv", sep = ",", header = TRUE)
nm_boxplot <- ggplot(nm_pref_data, aes(x= NM_Den, y = NM_Perc_Den, fill=NM_Den)) +
  geom_boxplot() + theme_classic() +
  labs (x="Den", y="Time in Den at New Moon (%)") +
  scale_fill_manual(values=c("#D8B365", "#5AB4AC")) + theme(legend.position = "none") +
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14))+ ylim(0,100) +scale_x_discrete(limits=c("Control", "Magnet"))
nm_boxplot
ggsave("newmoon.pdf", width=3, height=4, dpi=300)
```

```{r}
compare_means(NM_Perc_Den ~ NM_Den, data = nm_pref_data, paired = TRUE)
#NM_Perc_Den	Control	Magnet	0.7275585	0.73	0.73	ns	Wilcoxon
```


```{r}
fm_pref_data <- read.csv("2024_01_25 Full Moon Den Preference.csv", sep = ",", header = TRUE)
fm_boxplot <- ggplot(fm_pref_data, aes(x= FM_Den, y = FM_Perc_Den, fill=FM_Den)) +
  geom_boxplot() + theme_classic() +
  labs (x="Den", y="Time in Den at Full Moon (%)") +
  scale_fill_manual(values=c("#D8B365", "#5AB4AC")) + theme(legend.position = "none") +
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14))+ ylim(0,100) +scale_x_discrete(limits=c("Control", "Magnet")) 
fm_boxplot
ggsave("fullmoon.pdf", width=3, height=4, dpi=300)
```


```{r}
compare_means(FM_Perc_Den ~ FM_Den, data = fm_pref_data, paired = TRUE)
#FM_Perc_Den	Control	Magnet	0.03933871	0.039	0.039	*	Wilcoxon
```

################################################################################
Total Time in a Den Plotting and Statistical Analysis

```{r}
total_den_data <- read.csv("2024_01_25 Total Time in Den.csv", sep = ",", header = TRUE)
den_boxplot <- ggplot(total_den_data, aes(x= Moon_Cycle, y = Perc_In_Any_Den, fill=Moon_Cycle)) +
  geom_boxplot() + theme_classic() +
  labs (x="Lunar Phase", y="Total Time in a Den (%)") + scale_fill_manual(values=c("#c7eae5", "#f6e8c3")) + theme(legend.position = "none") +
  theme(axis.text = element_text(size=12), axis.title = element_text(size=14))+ ylim(0,100) +scale_x_discrete(limits=c("New Moon", "Full Moon"))
den_boxplot
ggsave("denboxplot.pdf", width=6, height=4, dpi=300)
```


```{r}
compare_means(Perc_In_Any_Den ~ Moon_Cycle, data = total_den_data, paired = FALSE)
```

################################################################################
Magnet Field Strength vs Position in the Tank PLotting

```{r}
magnet_data <- read.csv("2024_01_12 Magnet Strength.csv", sep = ",", header = TRUE)
magnet_data
magnet_tank <- ggplot(magnet_data, aes(x= Position, y = Corrected_Field_Strength)) +
  geom_line() + geom_point()+ theme_classic() +
  labs (x="Position in Tank (cm)", y="Field Strength (μT)")+ theme(axis.text = element_text(size=12), axis.title = element_text(size=14))+ ylim(0,321) + xlim(0,110) + scale_x_continuous(breaks = seq(0, 110, 10)) + scale_y_continuous(breaks = seq(0, 321, 50))
magnet_tank

ggsave("magnettank.pdf", width=6, height=4, dpi=300)
```

